Individual Differenes

### 2023-08-30 Misc/benchmarks.qmd
alternate model specifications - multinomial

>>>>

```{r}
source(here::here("Functions", "packages.R"))
test <- readRDS(here("data/e1_08-21-23.rds")) |>  filter(expMode2 == "Test") |>
  select(id,condit,bandInt,vb,vx,dist,sdist,bandType)

e1_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e1_testVxBand_RF_5k")))
new_data_grid=map_dfr(1, ~data.frame(unique(test[,c("id","condit","bandInt")])))

t1 <- system.time({
tidy_pred <- test |> add_predicted_draws(e1_vxBMM) |> mutate(.residual = vx - .prediction); 
predict_per_row <- tidy_pred |> group_by(.row) |> mean_hdi(.prediction,.residual)

})

m1 <- system.time({
  predict_vx <- test |> cbind(predict(e1_vxBMM, test)) |> mutate(resid=vx-Estimate)
}
)

cat(paste0("tidybayes preds = ",round(t1["elapsed"],2)," \n ",
           "predict.brmsfit = ",round(m1["elapsed"],2)))

predict_vx |> group_by(id,condit,bandInt) |> summarise(mean(vx),mean(Estimate),mean(resid))

```
tidybayes preds = 169.94 
predict.brmsfit = 10.27

### 2023-08-30 Misc/gam_discrim.qmd
alternate model specifications - multinomial

```{r}
pacman::p_load(tidyverse,lme4,emmeans,here,knitr,kableExtra,gt,mgcv,data.table)
e1 <- readRDS(here("data/e1_08-21-23.rds"))
source(here("Functions/Display_Functions.R"))

test <- e1 |> filter(expMode %in% c("test-Nf","test-train-nf"))
testAvg <- test %>% group_by(id, condit, vb, bandInt) %>%
  summarise(vx=mean(vx),dist=mean(dist)) %>% data.table()

>>>>

```{r}
pacman::p_load(tidyverse,lme4,emmeans,here,knitr,kableExtra,gt,mgcv,data.table)
source(here::here("Functions", "packages.R"))

test <- readRDS(here("data/e1_08-21-23.rds")) |> filter(expMode2 == "Test")  |>
  select(id,condit,expMode, trial, bandInt,vb,vx,dist,vy) |>   mutate(condit_dummy = ifelse(condit == "Varied", 1, 0))  |> data.table()

train <-readRDS(here("data/e1_08-21-23.rds")) |> filter(expMode2 == "Train" | expMode=="train-Nf")  |>
  select(id,condit,expMode,trial,bandInt,vb,vx,dist,vy) |> mutate(condit_dummy = ifelse(condit == "Varied", 1, 0))  |> data.table()

testAvg <- test %>% group_by(id, condit, vb, bandInt) %>%
  summarise(vx=mean(vx)) %>% data.table()

trainAvg <- train %>% group_by(id, condit, vb, bandInt) %>%
  summarise(vx=mean(vx)) %>% data.table()

est_gam <- function(k, train_data, test_data) 
{

  # define train and test data
  train_data <- copy(train_data)
  test_data <- copy(test_data)

  #--- train a model ---#  
  #trained_model <- gam(vx ~ condit + s(bandInt, k = k)+ s(id, bandInt,bs='re'), data = train_data)
  trained_model <- gam(dist~  s(bandInt,k=2) + s(vy,k=k) + s(id,bs="re"), data = train_data)

  #--- predict y for the train and test datasets ---#
  train_data[, y_hat := predict(trained_model, newdata = train_data)] 
  train_data[, type := "Train"]
  test_data[, y_hat := predict(trained_model, newdata = test_data)] 
  test_data[, type := "Test"]

  #--- combine before returning ---#
  return_data <- 
    rbind(train_data, test_data) %>%
    .[, num_knots := k]

sim_results <- 
  lapply(1:6, function(x) est_gam(x, train, test)) %>%
  rbindlist()

sim_results <- 
  lapply(c(1,2,3,5,10,20,50), function(x) est_gam(x, train, test)) %>%
  rbindlist()

# sim_results <- 
#   lapply(1:6, function(x) est_gam(x, trainAvg, testAvg)) %>%
#   rbindlist()

ggplot(sim_results) +
  geom_point(data = train, aes(y = dist, x = bandInt), color = "grey") +
  geom_line(aes(y = y_hat, x = vy, color = factor(num_knots))) +
  theme_bw()

ggplot(sim_results[num_knots  %in% 1:6, ],aes(fill=vb)) + 
   stat_summary(data=train,geom = "bar", position=position_dodge(), fun = mean) +
   geom_line(aes(y = y_hat, x = bandInt, color = factor(num_knots)))

summary <- 
  sim_results[, .(mse = mean((y_hat - dist)^2)), by = .(num_knots, type)]

ggplot(summary) +
  geom_line(aes(y = mse, x = num_knots, color = type)) +
  geom_point(aes(y = mse, x = num_knots, color = type)) +
  xlab("Number of knots") +
  ylab("MSE") +
  scale_color_discrete(name = "") +
  theme_bw()

#https://fromthebottomoftheheap.net/2021/02/02/random-effects-in-gams/
mod_gam1 = gam(vx ~ 1 + (bandInt, bs = "cr",k=4) + s(id, bandInt,bs='re'), data = test)

gy = gam(vx ~ s(vy,k=2), data = test)
summary(gy)
coef(gy)
gam.check(gy)

gy = gam(vx ~ s(vy,k=6), data = test)
summary(gy)
coef(gy)
gam.check(gy)

gy = gam(vx ~ s(bandInt,k=5) + s(vy), data = test)
summary(gy)
coef(gy)
gam.check(gy)

gy = gam(vx ~ s(bandInt,k=2) + s(vy) + s(id,bs="re"), data = test)
summary(gy)
coef(gy)
gam.check(gy)

g1=gam(vx ~ s(bandInt,k=1),data=train)
summary(g1)
g1$coefficients

g2=gam(vx ~ s(bandInt,k=2),data=trainAvg)
summary(g2)
g2$coefficients

g3=gam(vx ~ s(bandInt,k=4),data=train)
summary(g3)
g3$coefficients

k1=bam(condit_dummy ~ s(vx,k=12),data=train,family=binomial())
summary(k1)

k2=bam(bandInt ~ s(vx,k=5),data=train)
summary(k2)

g2=gam(vx ~ s(bandInt,k=6),data=trainAvg)

id1 <- test |> filter(id==1)

g2=gam(vx ~ 0 + s(bandInt,k=6),data=id1)
summary(g2)
g2$coefficients
kmeans(id1$vx,6)

kn=knnreg(vx ~ bandInt,data=test)
summary(kn)

g4 <- gam(vx ~ bandInt + s(id,bs="re"),data=test)
summary(g4)
g4$coefficients

g5 <- gam(vx ~ condit+bandInt + s(id,bs="re",k=4),data=test)
summary(g5)
g5$coefficients
coef(g5)

g6 <- gam(vx ~ condit+bandInt + s(id,bs="re")+ s(id,bandInt,bs="re",k=6),data=test,method = 'REML')
summary(g6)
g6$coefficients

hits = test |> filter(dist==0)

hits |> ggplot(aes(x=bandInt,y=vx)) + geom_point()

hg1 <- gam(vx ~ 0 + s(bandInt,k=2),data=hits)
summary(hg1)
hg1$coefficients
plot(hg1)

### 2023-08-30 Misc/indv_diff.qmd
alternate model specifications - multinomial

fitted(e1_vxBMM,re_formula=NA)
fitted(e1_vxBMM,re_formula=NULL) # matches epred

@@

 indvDraws <- left_join(random_effects, fixed_effects, by = join_by(".chain", ".iteration", ".draw")) |> 
  rename(bandInt_RF = bandInt,RF_Intercept=Intercept) |>
  right_join(new_data_grid, by = join_by("id")) |> 
  mutate(
    Slope = bandInt_RF+b_bandInt,
    Intercept= RF_Intercept + b_Intercept,
    estimate = (b_Intercept + RF_Intercept) + 
      (bandInt*(b_bandInt+bandInt_RF)) + 
      (condit_dummy*b_conditVaried)+
      (bandInt * condit_dummy) * `b_conditVaried:bandInt`,
    SlopeInt=bandInt_RF+b_bandInt + ((`b_conditVaried:bandInt` + bandInt_RF)*condit_dummy),
    SlopeInt2=bandInt_RF+b_bandInt + ((`b_conditVaried:bandInt` + b_bandInt)*condit_dummy)
  )

@@

  ggplot(indvSlopes, aes(x = Slope, color = condit)) + 
  geom_density() 
    ggplot(indvSlopes, aes(x = SlopeInt, color = condit)) + 
  geom_density() 
```

@@

{ indvDraws |> indv_model_plot(indvSlopes ,test, slopeVar=Slope, rank_variable=Slope,n_sbj=n_sbj,"max") + 
     ggtitle("Slope = bandInt_RF+b_bandInt") } /
 {ggplot(indvSlopes, aes(x = Slope, color = condit)) + geom_density() } /
#gridExtra::tableGrob( indvSlopes |> group_by(condit) |> summarise(Mean=mean(Slope),Sd=sd(Slope)) ) + 
  gt_temp(indvSlopes |> group_by(condit) |> summarise(Mean=mean(Slope),Sd=sd(Slope)) ) +
  plot_layout(heights=unit(c(16,6,2),c('cm'))) + plot_annotation("vx ~ bandInt + (1 + bandInt || gr(id, by = condit))  ")

>>>>

# fitted(e1_vxBMM,re_formula=NA)
# fitted(e1_vxBMM,re_formula=NULL) # matches epred

@@

 indvDraws <- left_join(random_effects, fixed_effects, by = join_by(".chain", ".iteration", ".draw")) |> 
  rename(bandInt_RF = bandInt,RF_Intercept=Intercept) |>
  right_join(new_data_grid, by = join_by("id")) |> 
  mutate(
    Slope = bandInt_RF+b_bandInt,
    Intercept= RF_Intercept + b_Intercept,
    estimate = (b_Intercept + RF_Intercept) + 
      (bandInt*(b_bandInt+bandInt_RF)) + 
      (condit_dummy*b_conditVaried)+
      (bandInt * condit_dummy) * `b_conditVaried:bandInt`,
    SlopeInt=bandInt_RF+b_bandInt + ((`b_conditVaried:bandInt` + bandInt_RF)*condit_dummy),
    SlopeInt2=bandInt_RF+b_bandInt + ((`b_conditVaried:bandInt`)*condit_dummy)
  )

@@

  ggplot(indvSlopes, aes(x = Slope, color = condit)) + 
  geom_density() 
    ggplot(indvSlopes, aes(x = SlopeInt2, color = condit)) + 
  geom_density() 
```

```{r fig.width=15, fig.height=18}

e1_vxBMM %>% as_draws_df(variable = c('^b_', 'sigma'), regex = TRUE) |>
  select(starts_with("b_"),sigma) |> 
  pivot_longer(
    everything(),
    names_to = 'key',
    values_to = 'value') %>% 
  ggplot(aes(x = value, fill = key)) +
  stat_halfeye(
    # plot the mode
    point_interval = mode_hdi,
    # plot HDIs be at both the 95% levels
    .width = c(.95),
    outline_bars = TRUE,
    normalize = 'panels',
    show.legend = FALSE
  ) +
  facet_wrap(~key, scales = 'free', ncol = 4) +
  ggtitle('Posterior distributions')

 indvDraws |> filter(id %in% 1:10) |> group_by(id,condit,.draw) |> summarise(Slope=mean(Slope)) |> 
   ggplot(aes(x=Slope,col=condit))+geom_density() + facet_wrap(~id)

  indvDraws  |> group_by(id,condit,.draw) |> summarise(Slope=mean(Slope)) |> 
   ggplot(aes(x=Slope,col=condit,fill=condit))+stat_halfeye()

indvDraws |> group_by(condit,.draw) |> summarise(Slope=mean(Slope)) |> 
   ggplot(aes(x=Slope,col=condit,fill=condit))+stat_halfeye()

 indvDraws |> group_by(id,condit) |> summarise(Slope=mean(Slope)) |> 
   ggplot(aes(x=Slope,col=condit,fill=condit))+geom_density()

  indvDraws |> group_by(id,condit,.draw) |> summarise(SlopeInt2=mean(SlopeInt2)) |> 
   ggplot(aes(x=SlopeInt2,col=condit,fill=condit))+geom_density()

  # pp_check(e1_vxBMM,
  #        type = 'stat_grouped',
  #        ndraws = 200,
  #        group = 'id',
  #        stat = 'mean')

@@

{ indvDraws |> indv_model_plot(indvSlopes ,test, slopeVar=Slope, rank_variable=Slope,n_sbj=n_sbj,"max") + 
     ggtitle("Slope = bandInt_RF+b_bandInt") } /
 {ggplot(indvSlopes, aes(x = Slope, color = condit)) + geom_density() } /
 gridExtra::tableGrob( indvSlopes |> group_by(condit) |> summarise(Mean=mean(Slope),Sd=sd(Slope)) ) + 
 # gt_temp(indvSlopes |> group_by(condit) |> summarise(Mean=mean(Slope),Sd=sd(Slope)) ) +
  plot_layout(heights=unit(c(16,6,2),c('cm'))) + plot_annotation("vx ~ bandInt + (1 + bandInt || gr(id, by = condit))  ")
